---
title: "A Study on Genre Multiplicity"
author: "Chujun He, Sichu Chang, Yining Hua"
output: html_document
---
```{r, include=FALSE}
library(tidyverse)
library(mdsr)
library(RMySQL)
```

```{r}
db <- dbConnect_scidb(dbname = "imdb")
newtable<-db %>%
  dbGetQuery("SELECT t.id, t.title, t.production_year,
             mi.info as info
             FROM title t
             Join movie_info as mi on mi.movie_id = t.id
             where info_type_id = 3
             having production_year IS NOT NULL;")
```

```{r}
glimpse(newtable)
#a smaller table, from year 1901-1999 
#Since some data is missing from 2018 to 2115, we filter out these years' data
film_before_2017 = newtable%>%
  filter(production_year<=2017)
glimpse(film_before_2017)
```

Q:Is is true that older movies tended to fit squarely in one genre, while more recent movies tend to span multiple genres?

```{r}
#Find each film has how many genres in each year from 1874 to 2017
film_genre = film_before_2017%>%
   group_by(id,production_year)%>%
   summarise(sum_genre = n_distinct(info))%>%
   arrange(production_year)
film_genre
```

```{r}
# we define the film with more than 3 genres as the film with multilple genres and 
# find out how many films has more than 3 genres in each year from 1874 to 2017
multi_genre_film = 
  film_genre %>% 
  filter(sum_genre >= 3) %>% #want the films with overlapping genres (more than 3)
  group_by(production_year) %>% 
  summarise(multi_num = n_distinct(id))

multi_genre_film
```


```{r}
#percentages of multi_genres film each year
join_total_multi = 
  total_genre %>% left_join(multi_genre_film, by ='production_year') %>%
  mutate(multi_num_new = ifelse(is.na(multi_num),0,multi_num)) %>%
  mutate(multi_percent = multi_num_new/total_num)

join_total_multi
```

```{r}
ggplot(join_total_multi,aes(x=production_year,y=multi_percent))+
  geom_line()+
  geom_smooth()+
  theme_minimal()
```

From the graph, we can see that the percentage of movies that have multi-genres increaseed from 1900 to 1940, and then it kept decreasing till 2000. 
Overall, it is true that movies made today tend to have overlapping genres than before. Also, t seems that the 1947 court decision that broke up the vertical integretion of studios has contributed to the decrease of overlapping genres.

**The possible causes for dramatic increase from 2003 might be:**
2003:Hollywood action film star Arnold Schwarzenegger was elected the 38th governor of California.
2004: Facebook launched.
2005: The first YouTube video was uploaded to YouTube.com in late spring.
2005: The action sequel Charlie's Angels: Full Throttle (2003), the first Sony Pictures film, was also the first feature film to be released on Blu-Ray Disc, a next generation, high-definition optical disc format. 
2006: The social networking and blogging site Twitter was founded and debuted.
2007: Netflix became the first subscription service to offer ad-free, legal video streaming of movies directly onto one’s computer screen.
(Retrieved from http://www.filmsite.org/2000sintro.html)

```{r}
#How many different genres appear in each year from 1874 to 2017 
info_type_year = film_before_2017%>%
  group_by(production_year)%>%
  summarise(sum = n_distinct(info))%>% #how many different genres each year
  arrange(production_year)
info_type_year
```

From the data we can see that the number of distinct genres increases by years.

```{r}
#A graph of genres overyears
genre_oyG = info_type_year%>%
  ggplot(aes(x = production_year,y = sum))+
    geom_line()
genre_oyG
```


```{r}
#A graph of total films made overyear
film_oyG = film_before_2017%>%
  group_by(production_year)%>%
  summarize(sum_films = n())%>%
  ggplot(aes(x = production_year,y = sum_films))+
    geom_line()
film_oyG
```

From the 2 graphs above we can see that the number of genre types had a big increase from 1890 to 1910. The total amount of films had an increase accordingly during this period of time, but it soon decreased in around 1910-1920, mighty because of WW1. The amount of films didn't reach the amount once reached in around 1903 untill around 1995. However WW1 didn't decrease the genres of films made, but increased them.

